---
title: "Cleaning and Transformation"
format: html
editor: visual
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(readr)
library(dplyr)
#library(writexl) write it to csv instead
```

```{r}

# Load the 2024/5 version of the ORDA survey data
orda_raw <- read_csv("data/Access Folk Survey_cleansed for ORDA_FINAL.csv")

data <- orda_raw

```

```{r store questions in txt, echo=FALSE}
# Extract column names
column_names <- colnames(data)

# Write column names to a text file
write.table(column_names, file = "./data/questions.txt", col.names = FALSE, row.names = FALSE, quote = FALSE)
rm(column_names)
```

## cannot create column Organiser

```{r , echo=FALSE}

your_data <- data
# Separate the column into multiple rows
your_data_long <- separate_rows(your_data, `Q21 - Have you organised non-folk music events as well? - Selected Choice`, sep = ",")

# Spread the data back into columns
your_data_wide <- spread(your_data_long, key = `Q21 - Have you organised non-folk music events as well? - Selected Choice`, value = `Q21 - Have you organised non-folk music events as well? - Selected Choice`)

your_data_wide$`<NA>`<- NULL
data<- your_data_wide
rm(your_data,your_data_long,your_data_wide)
```

Create a new column "Organiser" (Professional,Volunteer,Non-organiser) and remove questions 20 and 21 and all other columns created for that purpose

```{r Organiser, echo=FALSE,eval=FALSE}


data$Organiser <- ifelse(
  !is.na(data[, 76]) | !is.na(data[, 75]),   # Check if either column 76 or 75 has a value
  "Professional",
  ifelse(
    apply(data[, c(74, 77:80)], 1, function(x) any(!is.na(x))),  # Check if any of the other columns have a value
    "Volunteer",
    "Non-organiser"
  )
)

# Delete columns 74 to 80
data <- subset(data, select = -c(74:80))


# Check if all rows with "No" in column 20 have "Non-organiser" in column 74
result <- all(data[!is.na(data[, 20]) & data[, 20] == "No", 74] == "Non-organiser", na.rm = TRUE)
print(result)
data$`Q20: Have you ever been involved in organising folk singing events as an organiser, promoter, volunteer or in another way?`<- NULL

```

## Creating column: Ethnicity

```{r Ethnicity, echo=FALSE}

unique_values <- table(data$`Q53 - What best describes your ethnic group? - Selected Choice`)

# Create a new column 'Ethnicity' based on conditions
data <- data %>%
  mutate(
    Ethnicity = ifelse(
      is.na(`Q53 - What best describes your ethnic group? - Selected Choice`) | `Q53 - What best describes your ethnic group? - Selected Choice` == "Prefer not to say",
      "NA",
      ifelse(
        `Q53 - What best describes your ethnic group? - Selected Choice` %in% c(
          "White English/Welsh/Scottish/Northern Irish/British",
          "White Cornish"
        ),
        "White British",
        ifelse(
          `Q53 - What best describes your ethnic group? - Selected Choice` %in% c(
            "Any other white background, please describe",
            "White Gypsy or Traveller",
            "White Irish",
            "Any other ethnic group, please describe"
          ),
          "White Other",
          ifelse(
            `Q53 - What best describes your ethnic group? - Selected Choice` %in% c(
              "Any other Mixed/Multiple ethnic background, please describe",
              "White and Asian",
              "White and Black Caribbean"
            ),
            "Mixed/multiple",
            ifelse(
              `Q53 - What best describes your ethnic group? - Selected Choice` %in% c(
                "African",
                "Any other Black/African/Caribbean background, please describe",
                "Indian"
              ),
              "Non-White",
              "Unknown"
            )
          )
        )
      )
    )
  )

data$`Q53 - What best describes your ethnic group? - Selected Choice`<- NULL
data$Ethnicity <- as.factor(data$Ethnicity)
```

## Creating column: Gender

```{r Gender, echo=FALSE}

# Create a new column 'Gender' based on conditions
data <- data %>%
  mutate(
    Gender = case_when(
      `Q45 - What best describes your gender? - Selected Choice` == "Male" ~ "Male",
      `Q45 - What best describes your gender? - Selected Choice` == "Female" ~ "Female",
      TRUE ~ "Other"
    )
  )

data$Gender <- as.factor(data$Gender)

data$`Q45 - What best describes your gender? - Selected Choice`<-NULL
data$`Q45_6_TEXT - What best describes your gender? - In another way, please describe - Text`<-NULL

```

until here: data has: 801 obs and 57 var

## Creating column: Disability

```{r Disability, echo=FALSE}

data$Disability<- as.factor(data$`Q56 - Do you consider yourself to be d/Deaf or disabled, or do you have a long-term physical or mental health condition? (We will not ask you to identify any disability or health condition).`)

# Replace "Prefer not to say" with NA in your column
data <- data %>%
  mutate(Disability = case_when(
    Disability == "Prefer not to say" ~ NA_character_,
    TRUE ~ Disability
  ))

data$Disability<-as.factor(data$Disability)
data$`Q56 - Do you consider yourself to be d/Deaf or disabled, or do you have a long-term physical or mental health condition? (We will not ask you to identify any disability or health condition).`<- NULL
```

## Creating column: Sexuality

```{r Sexuality, echo=FALSE}
data$Sexuality<- data$`Q46 - How would you describe your sexual orientation? - Selected Choice`

data <- data %>%
  mutate(Sexuality = case_when(
    Sexuality == "Prefer not to say" ~ NA_character_,
    TRUE ~ Sexuality
  ))

data$Sexuality<-as.factor(data$Sexuality)
data$`Q46 - How would you describe your sexual orientation? - Selected Choice`<-NULL

```

## Creating column: Age

```{r Age, echo=FALSE}
data$Age<-as.factor(data$`Q44 - How old are you?`)

data$`Q44 - How old are you?`<-NULL

```

#### Summary for the Demographic Variables

```{r Summary, echo=FALSE}

summary(data$Age)
summary(data$Gender)

summary(data$Disability)
summary(data$Ethnicity)

summary(data$Organiser)

summary(data$Sexuality)
```

```{r , echo=FALSE,eval=FALSE}

# Education distribution
education_plot <- your_data %>%
  plot_ly(x = ~Q52, type = "bar", marker = list(color = ~Q52)) %>%
  layout(title = "Education Distribution", xaxis = list(title = "Education"), yaxis = list(title = "Count"))

# Social class distribution
class_plot <- data %>%
  plot_ly(x = ~`Q54: Which social class do you most identify with? - Selected Choice`, type = "bar", marker = list(color = ~`Q54: Which social class do you most identify with? - Selected Choice`)) %>%
  layout(title = "Social Class Distribution", xaxis = list(title = "Social Class"), yaxis = list(title = "Count"))

# Religion distribution
religion_plot <- your_data %>%
  plot_ly(x = ~Q67, type = "bar", marker = list(color = ~Q67)) %>%
  layout(title = "Religion Distribution", xaxis = list(title = "Religion"), yaxis = list(title = "Count"))

# Disability distribution check Q28

```

## Creating column: Political

```{r }
# Rename the column
colnames(data)[which(colnames(data) == "Q57 - If there was a general election tomorrow, which political party would you be most likely to support? - Selected Choice")] <- "Political_party"

# Group into 5 categories
data <- data %>%
  mutate(
    Political_party = case_when(
      Political_party %in% c("Labour Party", "Green Party", "Liberal Democrats", "Conservative Party") ~ Political_party,
      TRUE ~ "Other"
    ),
    Political_party = as.factor(Political_party)
  )


colnames(data)[which(colnames(data) == "Q58_1 - How politically engaged do you consider yourself to be?
0 = not politically engaged at all
10 = very politically engaged")] <- "Political_engagement"


data$Political_engagement <-as.factor(data$Political_engagement)

```

## data transformation and saving the data

```{r , echo=FALSE,warning=FALSE,message=FALSE}

colnames(data)[8]<- "Microphone"

colnames(data)[9]<- "Standing"

colnames(data)[10]<- "Instrument"

colnames(data)[11]<- "Word"

data$Microphone<-as.factor(data$Microphone)
data$Standing<-as.factor(data$Standing)
data$Instrument<-as.factor(data$Instrument)
data$Word<-as.factor(data$Word)




#change ehtnicity into three categories (4th: Unknown)
 

 data <- data %>%
  mutate(
    Ethnicity = case_when(
      Ethnicity %in% c("Mixed/multiple", "Non-White") ~ "Any Other Ethnicity",
      Ethnicity %in% c("NA", "Unknown") ~ "Unknown",
      TRUE ~ as.character(Ethnicity)
    ),
    Ethnicity = as.factor(Ethnicity)
  )

#Create new sexuality column called sexuality_cleand 
 
 # Create a new column for cleaned sexuality
data <- data %>%
  mutate(
    Sexuality_clean = case_when(
      is.na(Sexuality) ~ "NA",
      str_detect(Sexuality, regex("straight|heterosexual", ignore_case = TRUE)) ~ "Straight",
      str_detect(Sexuality, regex("bisexual", ignore_case = TRUE)) ~ "Bisexual",
      str_detect(Sexuality, regex("gay|lesbian", ignore_case = TRUE)) ~ "Gay/Lesbian",
      str_detect(Sexuality, regex("queer", ignore_case = TRUE)) ~ "Queer",
      str_detect(Sexuality, regex("questioning", ignore_case = TRUE)) ~ "Questioning",
      TRUE ~ "Other"
    )
  )




#save new data
write.csv(data, file ="./data/final_dataset.csv", row.names = FALSE)


```
